\chapter{Acknowledgments}
% 8. Acknowledgments{tc  \l 1 "8. Acknowledgments} 

A large number of people have contributed to the SEISAN system. From the British Geological Survey, Jim Bolton has spent several months cleaning up the software, putting in error checking and put in the new version of EPIMAP. Jane Exton has also been involved in several of the SEISAN database problems. A main contributor has been Barry Lienert who has spent several months at our institute to modify and adopt the HYP program, he has also written the complete HYP manual. Bent Ruud has contributed with the core of the AUTOPIC software and has helped out in practical aspects of installing it. Kuvvet Atakan has written the seismic hazard part. Ronald Arvidson has tested large parts of the system and done work on several programs and contributed with the modified version of FOCMEC. The help of Arne Sjursen has been essential for implementing X. Vunganai Midzi did the complete testing of version 6.0. Mario Villagran with programs, many suggestions and bug reports has interacted with the development of SEISAN. Ezra Twesigomwe, Berit Storheim, K. Atakan and Alice Walker have debugged the manuscript. 
Version 7: Bladimir Moreno has made the Windows graphics, made SEISAN run under Linux, written several other programs and has thus enormously contributed to this version. The hazard part has been updated by Kuvvet Atakan and Anibal Ojeda. Andrius Pacesa has implemented the FK routine. The programs have been tested and the manual been checked by Margaret Grandison, Waldo Taylor, Vunganai Midzi, Berit Storheim, Anne Lise Kj\ae rgaard, Anibal Ojeda, Ileana Boschini and Cecilie Langeland. 
Version 7.1: This version was tested by Anne Lise Kj\ae rgaard, Margaret Grandison and Vunganai Midzi. 
Version 7.2: Several contributions, including changes to MULPLT and implementing HYPO71, were made by Brian Baptie from the BGS. W\_EMAP is a new program that was provided by Fernando Carrilho. The CPLOT program was written by Susanne Lund Jensen from KMS. Susanne has also checked this version of the manual. 
Version 8.0: The graphics part on Unix has been improved thanks to Freya Cromarty and Frederik Tillmann. The on-site-inspection group at the CTBTO has financially supported the development of SEISCONF and JSEISAN, as well as the modification of other programs. JSEISAN and SEISCONF were written by Bladimir Moreno. A new version of W\_EMAP was provided by Fernando Carrilho. Mario Ordaz contributed the DEGTRA A4 software. Angel Rodriguez contributed the Perl reading routine and Terje Utheim the sfile Java program. Mathilde B\o ttger S\o rensen has revised the manual and tested the distribution. Dieter Stoll has provided information on how to compile on MacOSX and tested the software on Mac. RefTek has provided the rt\_seis program. 
Version 8.1: Rodrigo Luciano Pereira Canabrava made a major contribution to SEISAN by writing and implementing routines to read and write SEED data. Richard Luckett implemented the ISC location program. Mathilde B\o ttger S\o rensen wrote scripts for the new macroseismic part. Brian Baptie contributed tools to convert event data for use with travel time conversion programs. 
Version 8.3: Wayne Crawford has contributed to a number of programs and made valuable suggestions.
Version 9.0: Ruben Soares Lu\'is has made the interface routines for reading archives and contributed several Java programs.
Finally we will thank all the patient users who have suffered from the "bugs" and have given useful feedback. 


