
\subsection{IASPEI travel time software, program TTIM} 

This program can be used for calculating global travel times, see below for details on phases calculated. The program assumes that you have the travel time tables in the working directory or in DAT, see computer notes below on how to generate these file if not already there. The same files are also used by HYPOCENTER. 

After starting the program, the first two questions 'do you want xxxxx' relate to range summaries, etc., that are normally not required and can be answered with n(no) followed by ENTER. The program then asks 'Enter phases, one per line...' You can then enter a specific phase, or a keyword defined as follows:  


\begin{tabular}{lp{10cm}}
All & gives all phases \\
P & gives P-up, P, Pdiff, PKP, and Pkikp \\
P+ & gives P-up, P, Pdiff, PKP, Pkikp, PcP, Pp, Ppdiff, PPKP, 
PPKIKP, Sp, Spdiff, SPKP, and SPKIKP \\
S+ & gives S-up, S, Sdiff, SKS, Ss, Ssdiff, SSKS, Ps, Psdiff, and PSKS \\
basic & gives P+ and S+ as well as ScP, SKP, PKKP, SKKP, PP, and P'P' \\
\end{tabular}

Writing all individual phases, separate by ENTER, terminating the 
list with an additional ENTER.  The program will then enter a loop 
where phase times are calculated for new distances entered on request. 
The program is terminated for a particular distance by entering -1, 
and a new depth can be used, or the program can be terminated by entering -1 again. \newline
A special version of this program used in connection with MULPLT is IASP. \newline
In order to generate the earth model files \index{IASP91.HED}IASP91.HED 
and IASP91.TBL, first run program REMODL, then program SETBRN. The 
program REMODL has the earth model hardwired. Note: These binary files 
CANNOT be moved between platforms. They are included with SEISAN for 
each respective distribution. If lost, they must be regenerated on 
the same platform.  \newline
For more information about IASP91 programs, see HYPOCENTER manual by \textbf{B. Lienert}. 

%\textcolor{red}{pv-change: 
On at \index{64 bit computer} 64 bit 
computer the IASP files must be regenerated 
is you have the files from a 32 bit computer, with the programs \texttt{REMODL} and 
\texttt{SETBRN}. 


